Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging

### 2022-12-13 script/manuscript_MEE.Rmd
j edits inserted

output:
  word_document: default
  toc: no
  pdf_document: default
csl: "../citations/methods-in-ecology-and-evolution.csl"
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
always_allow_html: true
header-includes:
- \usepackage{endfloat}
- \usepackage{setspace}\doublespacing
- \usepackage{lineno}
- \linenumbers
---

@@

The inability to reliably identify plants to the level of species often leaves our understanding of ecosystem function and interactions wanting. Current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques. These approaches are untenable in light of the benefits offered by: species in several morphologically difficult genera which serve as bioindicators, preferred partners in ecological interactions, as well as an increasing lack of taxonomic experts (@hebert2003biological). Many genera, especially with the formalized advent of integrative taxonomy, have species which are well defined based upon ecological and behavioral rather than morphological properties, the identification of these taxa in degraded areas or without their mutualistic partners is fraught with difficulty. Hindering an understanding of the breadth of habitat which some species occupy, and the interactions they have with other species.

  The identification of many plant species to terminal taxon is an essential component of nearly all land management programs, where many species in the same genus  (e.g. Sagebrush - *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) serve as bioindicators (respectively for 'rangelands', streams, and wetlands), as well as in academic research (@Gage2013HistoricalRO, AIM). This endeavour is often mired by lack of diagnostic characters (e.g. flowers, fruits, roots or combinations thereof),  and increasingly the description of cryptic species (@janzen2017nuclear, @oliver2009cryptic). Solutions to this problem are wanting, certain programmes have relied increasingly upon revisiting field sites to identify material using morphological or chemical approaches, whereas academic research has often used high copy number plastid genes as barcodes (Rosentreter et al. 2021, MORE MORE). However, both approaches have significant downsides, the former resource intensive at the landscape scale - and often does not work, while the latter seldom works due to a lack of variability in the currently available barcodes (@liu2014identification).

@@

@@

@@

  We apply these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's is consistent across direct observations and the pollen record, an incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence). The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble). Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous populations declines, while simultaneously serving as the most effective pollinators in temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient). *Heterogeneous montane ecosystems simultaneously represent one of the most ecologically resilient areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst experiencing the most abrupt changes in mean annual temperature (CITE). An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if these refugial ecosystemn services are to be utilized while maintaining their current species richness.*

@@

A custom Kraken2 database was created by downloading representative species of each genus indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved). These sequences were processed in the same manner as our novel sequences before being placed into the database. The Kraken2 database was built using default parameters. Kraken2 was run on sequences using default parameters (*APPENDIX 5*). 
  Following Kraken2, Bracken was used to classify sequences to terminal taxa (@lu2017bracken). Results from both Kraken2 and Bracken, results were reclassified manually to identify terminal taxa. For example, when only a single species of a genus was known in the study area, but our database used a representative of another taxon in the genus, this species was coded as the result. The re-coding of sequences from another representative species for the genus to the sole RMBL representative allowed the identification of XX & % more species.

>>>>

output:
  pdf_document: default
  toc: no
  word_document: default
csl: "../citations/methods-in-ecology-and-evolution.csl"
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
always_allow_html: true
header-includes:
- \usepackage{endfloat}
- \usepackage{setspace}\doublespacing
- \usepackage{lineno}
- \linenumbers
---

@@

*< UTILITY OF MASS IDENTIFICATION OF MATERIAL - FRIENDLY STAGE SET >*

*< ISSUES AND COSTS WITH IDENTIFYING MATERIAL TO SPECIES >* many spp defined by interactions (<;

The inability to reliably identify plants down to species can limit our understanding of ecosystem function and interactions. This is especially true for genera where the species are well defined based upon ecological and behavioral rather than morphological properties, and hence can serve as key habitat bioindicators(e.g. different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO). The lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and the interactions they have with other species (CITE). Current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques.

  The identification to species is often mired by lack of diagnostic characters (e.g. flowers, fruits, roots or combinations thereof), an increasing lack of taxonomic experts (@hebert2003biological) and increasingly the description of cryptic species (@janzen2017nuclear, @oliver2009cryptic). And revisiting field sites to identify material using morphological or chemical approaches, can be resource intensive and often does not work.

  *< BARCODING HAS ITS LIMITS CURRENTLY BUT THERE IS POTENTIAL >*

@@

  *<  WHAT NOVEL APPROACES IN ADDITION TO A353 ?>*

@@

  *< THIS P -- WHAT WE DO TO TEST THIS OUT >*

@@

  To test these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's is consistent across direct observations and the pollen record, an incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence). The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble). Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous populations declines, while simultaneously serving as the most effective pollinators in temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient). *Heterogeneous montane ecosystems represent one of the most ecologically resilient areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the most abrupt changes in mean annual temperature (CITE). An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if these refugial ecosystemn services are to be utilized while maintaining their current species richness.*

@@

A custom Kraken2 database was created by downloading representative species of each genus indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved). These sequences were processed in the same manner as our novel sequences . The Kraken2 database was built using default parameters. Kraken2 was run on sequences using default parameters (*APPENDIX 5*). 
  Following Kraken2, Bracken was used to classify sequences to terminal taxa (@lu2017bracken). Results from both Kraken2 and Bracken, results were reclassified manually to identify terminal taxa. For example, when only a single species of a genus was known in the study area, but our database used a representative of another taxon in the genus, this species was coded as the result. The re-coding of sequences from another representative species for the genus to the sole RMBL representative allowed the identification of XX & % more species.
